---
title: "environmentalLayers.Rmd"
output: html_document
---


```{r}
require(raster)
require(terra)
require(sf)
require(rgdal)
require(tidyverse)
require(rgeos)
require(gridExtra)
require(factoextra) 
```


```{r}
## Importing biome limits ---------------------------
amaz = shapefile('./data/amazbioma.shp')
amaz = spTransform(amaz, CRS("+proj=longlat +datum=WGS84"))
```


```{r}
## Importing raster files ---------------------------

## u-speed ---------------------------
uspeed = abs(raster('./data/uspeed.tif'))
uspeed = setMinMax(uspeed)
ref = abs(uspeed)

## v-speed ---------------------------
vspeed = raster('./data/vspeed.tif') %>% crop(ref)
vspeed = abs(vspeed)
vspeed = raster(vals=values(vspeed),
                ext=extent(ref), 
                nrows=dim(ref)[1],
                ncols=dim(ref)[2])
vspeed = setMinMax(vspeed)

## Lightning ---------------------------
lightning = raster('./data/lightning.tif') %>% crop(ref)
lightning = raster(vals=values(lightning),
                   ext=extent(ref), 
                   nrows=dim(ref)[1],
                   ncols=dim(ref)[2])
lightning = setMinMax(lightning)

## Days with more than 20 mm ---------------------------
days20 = raster('./data/days20.tif') %>% crop(ref)
days20 = raster(vals=values(days20),
                ext=extent(ref), 
                nrows=dim(ref)[1],
                ncols=dim(ref)[2])
days20 = setMinMax(days20)

## elevation ---------------------------
elevation = raster('./data/elevation.tif') %>% crop(ref)
elevation = raster(vals=values(elevation),
              ext=extent(ref), 
              nrows=dim(ref)[1],
              ncols=dim(ref)[2])

elevation[elevation <= 0] = NA
elevation[elevation >= 1000] = NA
elevation = setMinMax(elevation)

## Potential Evapotranspiration ---------------------------
pet = raster('./data/pet.tif') %>% crop(ref)
pet = raster(vals=values(pet),
             ext=extent(ref), 
             nrows=dim(ref)[1],
             ncols=dim(ref)[2])
pet = setMinMax(pet)

## FAPAR ---------------------------
# fapar = raster('./data/fapar.tif') %>% crop(ref)
# summary(getValues(fapar))
# tm_shape(fapar) + tm_raster(n = 10)
# 
# fill.na <- function(x) {
#   center = 0.5 + (width*width/2) 
#   if( is.na(x)[center] ) {
#     return( median(x, na.rm=TRUE) )
#   } else {
#     return( x[center] )
#   }
# } 
# 
# fapar2 = faparteste
# before = summary(getValues(fapar2))[7]
# print(before)
# for(i in seq(1, 10, 1)){
#   width = 9
#   fapar2 = focal(fapar2, w = matrix(1,width,width), fun = fill.na, pad = TRUE, na.rm = FALSE) %>% crop(ref)
#   after = summary(getValues(fapar2))[7]
#   print(after)
# }
# 
# fapar2 = raster(vals=values(fapar2),
#                 ext=extent(ref), 
#                 nrows=dim(ref)[1],
#                 ncols=dim(ref)[2])
# fapar2 = setMinMax(fapar2)
# 
# writeRaster(fapar2, "./data/faparProcessed.tif", overwrite=TRUE)
# 
# tm_shape(fapar2) + tm_raster(n = 10)

fapar = raster('./data/fapar.tif') %>% crop(ref)

fapar = raster(vals=values(fapar),
               ext=extent(ref), 
               nrows=dim(ref)[1],
               ncols=dim(ref)[2])
fapar = setMinMax(fapar)


fapar2 = raster('./data/faparProcessed.tif') %>% crop(ref)
fapar2 = raster(vals=values(fapar2),
                ext=extent(ref), 
                nrows=dim(ref)[1],
                ncols=dim(ref)[2])
fapar2 = setMinMax(fapar2)


## Clear days ---------------------------
clearDays = raster('./data/clearDays.tif') %>% crop(ref)
clearDays = raster(vals=values(clearDays),
                   ext=extent(ref), 
                   nrows=dim(ref)[1],
                   ncols=dim(ref)[2])
clearDays = setMinMax(clearDays)

## Month with less than 100 mm ---------------------------
# month100 = raster('./data/month100.tif') %>% crop(ref)
# month100 = raster(vals=values(month100),
#                   ext=extent(ref), 
#                   nrows=dim(ref)[1],
#                   ncols=dim(ref)[2])
# month100 = setMinMax(month100)

## Annual precipitation ---------------------------
pannual = raster('./data/pannual.tif') %>% crop(ref)
pannual = raster(vals=values(pannual),
                 ext=extent(ref), 
                 nrows=dim(ref)[1],
                 ncols=dim(ref)[2])
pannual = setMinMax(pannual)

## Precipitation of the driest month ---------------------------
# pdriest = raster('./data/pdriest.tif') %>% crop(ref)
# pdriest = raster(vals=values(pdriest),
#                  ext=extent(ref), 
#                  nrows=dim(ref)[1],
#                  ncols=dim(ref)[2])
# pdriest = setMinMax(pdriest)

## Precipitation of the wettest month ---------------------------
pwettest = raster('./data/pwettest.tif') %>% crop(ref)
pwettest = raster(vals=values(pwettest),
                  ext=extent(ref), 
                  nrows=dim(ref)[1],
                  ncols=dim(ref)[2])
pwettest = setMinMax(pwettest)

## Precipitation seasonality ---------------------------
pseason = raster('./data/pseason.tif') %>% crop(ref)
pseason = raster(vals=values(pseason),
                 ext=extent(ref), 
                 nrows=dim(ref)[1],
                 ncols=dim(ref)[2])
pseason = setMinMax(pseason)

## Temperature seasonality ---------------------------
tseason = raster('./data/tseason.tif') %>% crop(ref)
tseason = raster(vals=values(tseason),
                 ext=extent(ref), 
                 nrows=dim(ref)[1],
                 ncols=dim(ref)[2])
tseason = setMinMax(tseason)

## Annual temperature ---------------------------
tannual = raster('./data/tannual.tif') %>% crop(ref)
tannual = raster(vals=values(tannual),
                 ext=extent(ref), 
                 nrows=dim(ref)[1],
                 ncols=dim(ref)[2])
tannual = setMinMax(tannual)

## Maximum temperature ---------------------------
tmax = raster('./data/tmax.tif') %>% crop(ref)
tmax = raster(vals=values(tmax),
              ext=extent(ref), 
              nrows=dim(ref)[1],
              ncols=dim(ref)[2])
tmax = setMinMax(tmax)

## Clay content ---------------------------
clayContent = raster('./data/clayContent30m.tif') %>% crop(ref)
clayContent = raster(vals=values(clayContent),
                     ext=extent(ref), 
                     nrows=dim(ref)[1],
                     ncols=dim(ref)[2])
clayContent = setMinMax(clayContent)
clayContent[clayContent <= 0] = NA  

## Water content ---------------------------
waterContent = raster('./data/waterContent30m.tif') %>% crop(ref)
waterContent = raster(vals=values(waterContent),
                      ext=extent(ref), 
                      nrows=dim(ref)[1],
                      ncols=dim(ref)[2])
waterContent = setMinMax(waterContent)
waterContent[waterContent <= 0] = NA  
```

```{r}
## Stack das vari치veis ambientais ---------------------------
layers = stack(uspeed, 
               vspeed)

```

## Stack das vari치veis ambientais ---------------------------
layers = stack(fapar, 
               elevation, 
               uspeed, 
               vspeed, 
               clearDays, 
               days20, 
               lightning, 
               #month100, 
               pannual, 
               #pdriest, 
               pet, 
               pseason, 
               pwettest, 
               tannual, 
               tseason, 
               tmax, 
               clayContent, 
               waterContent)

layers2estimate = stack(fapar2,
                        elevation,
                        uspeed,
                        vspeed,
                        clearDays,
                        days20,
                        lightning,
                        #month100,
                        pannual,
                        #pdriest,
                        pet,
                        pseason,
                        pwettest,
                        tannual,
                        tseason,
                        tmax,
                        clayContent,
                        waterContent)

layers2maxent = stack(#fapar2, 
                        elevation, 
                        #uspeed, 
                        #vspeed, 
                        clearDays, 
                        days20, 
                        lightning, 
                        #month100, 
                        pannual, 
                        #pdriest, 
                        pet, 
                        pseason, 
                        pwettest, 
                        tannual, 
                        tseason, 
                        tmax, 
                        clayContent, 
                        waterContent)

# save(layers2estimate, 
#     file = 'C:/Users/User/Documents/RStudio/secrets-robson/layers2estimate.Rdata')

```{r}
# atualiza nome das camadas
names(layers) = c("uspeed", 
                  "vspeed")
```

names(layers) = c("fapar", 
                  "elevation", 
                  "uspeed", 
                  "vspeed", 
                  "clearDays", 
                  "days20",
                  "lightning",
                  #"month100",
                  "pannual",
                  #"pdriest",
                  "pet",
                  "pseason",
                  "pwettest",
                  "tannual",
                  "tseason",
                  "tmax",
                  "clayContent",
                  "waterContent")

names(layers2estimate) = c("fapar",
                           "elevation",
                           "uspeed",
                           "vspeed",
                           "clearDays",
                           "days20",
                           "lightning",
                           #"month100",
                           "pannual",
                           #"pdriest",
                           "pet",
                           "pseason",
                           "pwettest",
                           "tannual",
                           "tseason",
                           "tmax",
                           "clayContent",
                           "waterContent")

names(layers2maxent) = c(#"fapar", 
                           "elevation",
                           #"uspeed", 
                           #"vspeed",
                           "clearDays",
                           "days20",
                           "lightning",
                           #"month100",
                           "pannual",
                           #"pdriest",
                           "pet",
                           "pseason",
                           "pwettest",
                           "tannual",
                           "tseason",
                           "tmax",
                           "clayContent",
                           "waterContent")

```{r}
## Open sites coordinates ---------------------------
sitioCoord = read.csv('data/siteForestPlots_en.csv')
coordinates(sitioCoord) <- ~Longitude+Latitude
proj4string(sitioCoord) <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
```


```{r}
## Cliping stack by maximum height location ------------------------
explanatoryVariables = raster::extract(layers, sitioCoord)  # extrai valor das vari치veis ambientais para cada altura m치xima mapeada
sitioCoord = cbind(sitioCoord, explanatoryVariables) 

```
